---
title: "7_correlations_with_metadata"
output: html_document
date: "2023-05-03"
---

#T-tests
```{r Check if there are actual differences Antibiotic all}
library(repmod)
library(tidyverse)
library(readxl)
metadata <- read_excel("Documents/Lola/CIPF_2018/Rats_HE_CCL4model/metadata.xlsx")
View(metadata)

metadata<-data.frame(metadata)

isVariableParametric <- function(data, variable) {
  # Perform Shapiro-Wilk test for normality
  normality_test <- shapiro.test(data[[variable]])
  
  # Perform Levene's test for homoscedasticity
  homoscedasticity_test <- car::leveneTest(data[[variable]], data$Antibiotic)
  
  # Check the p-values of both tests
  normality_p_value <- normality_test$p.value
  homoscedasticity_p_value <- homoscedasticity_test$Pr[1]
  
  # Check if the variable is parametric based on the test results
  is_parametric <- normality_p_value > 0.05 && homoscedasticity_p_value > 0.05
  
  return(is_parametric)
}

analysisSummary <- metadata %>%
    subset(select=-c(SampleID, Group, Week, Batch, Treated_CCL, FisabioID)) %>% 
    gather(key = variable, value = value, -Antibiotic) %>%
    group_by(variable, Antibiotic) %>% na.omit() %>%
    summarise(value=list(value)) %>%
    spread(Antibiotic, value) %>% 
    group_by(variable)

analysisSummary <- analysisSummary %>%
    mutate(parametric = isVariableParametric(metadata, variable))

analysisSummary <-analysisSummary %>% mutate(
    # Save the name of the test for convenient reference later
    test = case_when(
        parametric ~ "Welch's t test", TRUE ~ "Wilcoxon test"
    ),
    # Run the t-test for parametric variables and Wilcoxon signed rank test for non-parametric ones, save the p-value
    p_value = case_when(
        parametric ~ t.test(unlist(Yes), unlist(No))$p.value,
            TRUE ~ wilcox.test(unlist(Yes), unlist(No), paired=FALSE, exact=FALSE)$p.value
    ),
    # Run the test again, but now save the effect size
    statistic = case_when(
        parametric~ t.test(unlist(Yes), unlist(No))$statistic,
            TRUE ~ wilcox.test(unlist(Yes), unlist(No), paired=FALSE, exact=FALSE)$statistic
    ),
)

analysisSummary_ok<-analysisSummary %>% data.frame() %>%
    subset(select=-c(Yes, No))
  
make_csv_table(analysisSummary_ok, "~/Documents/Lola/CIPF_2018/Rats_HE_CCL4model/output/Tables/comparisons_antibiotic", info = "comparisons_antibiotic")
```

```{r Check if there are actual differences Treated_CCL all}
analysisSummary_T <- metadata %>%
    subset(select=-c(SampleID, Group, Week, Batch, Antibiotic, FisabioID)) %>% 
    gather(key = variable, value = value, -Treated_CCL) %>%
    group_by(variable, Treated_CCL) %>% na.omit() %>%
    summarise(value=list(value)) %>%
    spread(Treated_CCL, value) %>% 
    group_by(variable)

analysisSummary_T <- analysisSummary_T %>%
    mutate(parametric = isVariableParametric(metadata, variable))

analysisSummary_T <-analysisSummary_T %>% mutate(
    # Save the name of the test for convenient reference later
    test = case_when(
        parametric ~ "Welch's t test", TRUE ~ "Wilcoxon test"
    ),
    # Run the t-test for parametric variables and Wilcoxon signed rank test for non-parametric ones, save the p-value
    p_value = case_when(
        parametric ~ t.test(unlist(Yes), unlist(No))$p.value,
            TRUE ~ wilcox.test(unlist(Yes), unlist(No), paired=FALSE, exact=FALSE)$p.value
    ),
    # Run the test again, but now save the effect size
    statistic = case_when(
        parametric~ t.test(unlist(Yes), unlist(No))$statistic,
            TRUE ~ wilcox.test(unlist(Yes), unlist(No), paired=FALSE, exact=FALSE)$statistic
    ),
)

analysisSummary_Tok<-analysisSummary_T %>% data.frame() %>%
    subset(select=-c(Yes, No))
make_csv_table(analysisSummary_Tok, "~/Documents/Lola/CIPF_2018/Rats_HE_CCL4model/output/Tables/comparisons_Treated_CCL", info = "comparisons_antibiotic")
```

```{r Check if there are actual differences treated ccl vs ccl4+rif}
newmetadata2 <- metadata %>%
    group_by(Treated_CCL) %>%
    filter(!any(Treated_CCL == "No"))

analysisSummary_A_noccl <- newmetadata2 %>% 
    subset(select=-c(SampleID, Group, Week, Batch, Treated_CCL, FisabioID, GABAa5)) %>%
    gather(key = variable, value = value, -Antibiotic) %>%
    group_by(variable, Antibiotic) %>% na.omit() %>%
    summarise(value=list(value)) %>%
    spread(Antibiotic, value) %>% 
    group_by(variable)

analysisSummary_A_noccl <- analysisSummary_A_noccl %>%
    mutate(parametric = isVariableParametric(metadata, variable))

analysisSummary_A_noccl <-analysisSummary_A_noccl %>% mutate(
    # Save the name of the test for convenient reference later
    test = case_when(
        parametric ~ "Welch's t test", TRUE ~ "Wilcoxon test"
    ),
    # Run the t-test for parametric variables and Wilcoxon signed rank test for non-parametric ones, save the p-value
    p_value = case_when(
        parametric ~ t.test(unlist(Yes), unlist(No))$p.value,
            TRUE ~ wilcox.test(unlist(Yes), unlist(No), paired=FALSE, exact=FALSE)$p.value
    ),
    # Run the test again, but now save the effect size
    statistic = case_when(
        parametric~ t.test(unlist(Yes), unlist(No))$statistic,
            TRUE ~ wilcox.test(unlist(Yes), unlist(No), paired=FALSE, exact=FALSE)$statistic
    ),
)

analysisSummary_A_noccl_ok<-analysisSummary_A_noccl %>% data.frame() %>%
    subset(select=-c(Yes, No))
make_csv_table(analysisSummary_A_noccl_ok, "~/Documents/Lola/CIPF_2018/Rats_HE_CCL4model/output/Tables/analysisSummary_A_noccl_ok", info = "analysisSummary_A_noccl_ok")
```

```{r Check if there are actual differences ctr vs treated ccl}
newmetadata <- metadata %>%
    group_by(Antibiotic) %>%
    filter(!any(Antibiotic == "Yes"))%>%
    subset(select=-c(SampleID, Group, Week, Batch, Antibiotic, FisabioID, CCL5)) 

analysisSummary_T_norif <- newmetadata %>% 
    gather(key = variable, value = value, -Treated_CCL) %>%
    group_by(variable, Treated_CCL) %>% na.omit() %>%
    summarise(value=list(value)) %>%
    spread(Treated_CCL, value) %>% 
    group_by(variable)

analysisSummary_T_norif <- analysisSummary_T_norif %>%
    mutate(parametric = isVariableParametric(metadata, variable))

analysisSummary_T_norif <-analysisSummary_T_norif %>% mutate(
    # Save the name of the test for convenient reference later
    test = case_when(
        parametric ~ "Welch's t test", TRUE ~ "Wilcoxon test"
    ),
    # Run the t-test for parametric variables and Wilcoxon signed rank test for non-parametric ones, save the p-value
    p_value = case_when(
        parametric ~ t.test(unlist(Yes), unlist(No))$p.value,
            TRUE ~ wilcox.test(unlist(Yes), unlist(No), paired=FALSE, exact=FALSE)$p.value
    ),
    # Run the test again, but now save the effect size
    statistic = case_when(
        parametric~ t.test(unlist(Yes), unlist(No))$statistic,
            TRUE ~ wilcox.test(unlist(Yes), unlist(No), paired=FALSE, exact=FALSE)$statistic
    ),
)

analysisSummary_T_norif_ok<-analysisSummary_T_norif %>% data.frame() %>%
    subset(select=-c(Yes, No))
make_csv_table(analysisSummary_T_norif_ok, "~/Documents/Lola/CIPF_2018/Rats_HE_CCL4model/output/Tables/comparisons_Treatment_norif", info = "comparisons_Treatment_norif")
```

#Correlations
```{r Heatmap 1}
library(microbiome) # Load libraries
library(phyloseq)
library(dplyr)
library(reshape2)
library(knitr)

#If you wanted to select taxa and group
pseq2 <- ps_0 %>%
         subset_taxa(Phylum == "Bacteroidetes") %>%
         subset_samples(group == "LGG")


# Z transformed abundance data
pseqz <- microbiome::transform(ps_0, "Z")

# Load example data 
ps_3 <- tax_glom(ps_0, taxrank = 'Genus', NArm = FALSE) 
otu <- as.matrix(otu_table(pseqz))

# Define data sets to cross-correlate
metadata1 <- subset(metadata, select = -c(SampleID, Group, Antibiotic, Treated_CCL, FisabioID, Week, Batch))
metadata1 <- as.matrix(metadata1) 
rownames(metadata1) <- (metadata$SampleID)
x <- log10(t(otu))
y <- as.matrix(metadata1) 

# Cross correlate data sets
correlations <- associate(x, y, method = "spearman", mode = "matrix", p.adj.threshold = 0.05, n.signif = 1)

# Or, alternatively, the same output is also available in a handy table format
correlation.table <- associate(x, y, method = "spearman", mode = "table", p.adj.threshold = 0.05, n.signif = 1)

kable(head(correlation.table))
p <- heat(correlation.table, "X1", "X2", 
          fill = "Correlation", 
          star = "p.adj", 
          p.adj.threshold = 0.05) 

p + theme(text=element_text(size=10), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          legend.key.size = unit(1.3, "cm"))
```

```{r Heatmap 2}
# Z transformed abundance data
pseqz <- microbiome::transform(ps_0, "Z")
otu <- as.matrix(otu_table(pseqz))

# Define data sets to cross-correlate
metadata1 <- subset(metadata, select = -c(SampleID, Group, Antibiotic, Treated_CCL, FisabioID, Week, Batch))
metadata1 <- as.matrix(metadata1) 
rownames(metadata1) <- (metadata$SampleID)
x <- as.data.frame(t(otu))
y <- as.data.frame(metadata1) 
# Set black-and-white theme
library(ggplot2)
theme_set(theme_bw())

# Pick only the correlations with q<0.05
# Note: this will leave other cells empty
library(dplyr)
correlation.table <- associate(x, y, method = "spearman", mode = "table", p.adj.threshold = 0.95, n.signif = 1)
subtable <- filter(correlation.table, p.adj < 0.6)

# Arrange the figure
p <- ggplot(subtable, aes(x = X1, y = X2, fill = Correlation))
p <- p + geom_tile() 
p <- p + scale_fill_gradientn("Correlation", 
                              breaks = seq(from = -1, to = 1, by = 0.2), 
                              colours = c("darkblue", "blue", "white", "red", "darkred"), 
                              limits = c(-1,1)) 

# Polish texts
p <- p + theme(axis.text.x=element_text(angle = 90, hjust=1, face = "italic"),
               axis.text.y=element_text(size = 8))
p <- p + xlab("") + ylab("")

# Mark the most significant cells with stars
p <- p + geom_text(data = subset(correlation.table, p.adj < 0.1), 
                   aes(x = X1, y = X2, label = "+"), col = "white", size = 5)

# Plot
print(p)
```

###Correlations SCFAs
```{r correlation plot 1}
newmetadata <- metadata %>%
    subset(select=c(SCFA_AA, SCFA_BA, SCFA_PA, SCFA_CA, SCFA_VA, Group)) %>% data.frame()
newmetadata %>% 
    mutate(Group = factor(Group, 
                          levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif"))) %>% ggpairs(.,
                    title = "SCFAs per group", 
                    mapping = ggplot2::aes(colour=Group), 
                    lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.1), discrete = "points"), 
                    diag = list(discrete="barDiag", continuous = wrap("densityDiag", alpha=0.5 )), 
                    upper = list(combo = wrap("box_no_facet", alpha=0.5),continuous = wrap("cor", size=4))) + 
    theme(panel.grid.major = element_blank())
```

###Correlations some receptors
```{r correlation plot 2}
library(GGally)
newmetadata <- metadata %>%
    subset(select=c(GLAST, GLT1, TNFR1, NR2B, NR2A, NR1, GLUA1, GLUA2, Group)) %>% data.frame()
newmetadata %>% 
    mutate(Group = factor(Group, 
                          levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif"))) %>% ggpairs(.,
           title = "Receptor correlations per group", 
           mapping = ggplot2::aes(colour=Group), 
           lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.1), discrete = "points", combo = "blank"), 
           diag = list(discrete="barDiag", continuous = wrap("densityDiag", alpha=0.5 )), 
           upper = list(combo = wrap("box_no_facet", alpha=0.5), continuous = wrap("cor", size=4))) + 
    theme(panel.grid.major = element_blank())    # remove gridlines  
```

###Correlations cognition tests
```{r correlation plot 2}
library(GGally)
newmetadata <- metadata %>%
    subset(select=c(Ammonia, Vertcal_Counts, Stereotipic_Counts, Average_velocity, Ymaze_trialstolearn, Group)) %>% data.frame()
newmetadata %>% 
    mutate(Group = factor(Group, 
                          levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif"))) %>% 
  ggpairs(., title = "Cognitive tests",
          lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.1), discrete = "points", combo = "blank"), 
          diag = list(discrete="barDiag", continuous = wrap("densityDiag", alpha=0.5 )), 
          upper = list(combo = wrap("box_no_facet", alpha=0.5), continuous = wrap("cor", size=4))) + 
    theme(panel.grid.major = element_blank())    # remove gridlines  
```

###Correlations other metabolites and all boxplots
```{r correlation plot 2}
library(GGally)
newmetadata <- metadata %>%
    subset(select=c(GAT1, GAT3, OLM, IL_6, IL1_b, TNF_a, IFN_gamma, IL_15, IL_17, IL_10, IL_4, TGF_b, CCL2, CCL5, CCL20, CX3CL1, CCR2, CCR5, CX3CR1, Occludine, Learning_index_Rmaze)) %>% data.frame()
ggcorr(newmetadata, label = TRUE, label_alpha = TRUE, label_size = 3, angle = 310, hjust = 0.97, low = 'navy', high = 'orange')
ggcorr(newmetadata, label = TRUE, label_alpha = TRUE, label_size = 3, angle = 310, hjust = 0.97)
#all metadata
ggcorr(metadata, label = TRUE, label_alpha = TRUE, label_size = 3, angle = 310, hjust = 0.97)

analysisSummary <- metadata %>% 
    mutate(Group = factor(Group, 
                          levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif"))) %>%
    subset(select=-c(SampleID, Week, Batch, Treated_CCL, FisabioID, Antibiotic)) %>%
    gather(key = variable, value = value, -Group) %>%
    group_by(variable, Group) %>% na.omit() 
p <- ggplot(data = analysisSummary, aes(x=Group, y=value)) + 
    geom_boxplot(aes(fill=Group)) + facet_wrap( ~ variable, scales="free")  +
    theme(axis.text.x = element_text(size = 13), axis.title.y = element_text(size = 13), legend.text = element_text(size = 15))
```

#Shared core with venn diagram
Sudarshan Shetty, Leo Lahti, et al.
In may instances one would wish to compare core taxa that are shared between multiple groups. Here, we demonstrate how this can be achieved by microbiome and eulerr. The test data is stored in the microbiomeutilities R package and the original source of data is Zackular et al., 2014: The Gut Microbiome Modulates Colon Tumorigenesis.

devtools::install_github('microsud/microbiomeutilities')

```{r paquetes para diagrama core}
library(eulerr)
library(microbiome)
library(microbiomeutilities)
```

```{r Datos y modificaciones}
pseq5 <- microbiome::transform(ps_0, "identity")
# simple way to count number of samples in each group
table(meta(pseq5)$Group, useNA = "always")

# convert to relative abundances
 pseq.rel5 <- pseq5 %>%
   microbiome::transform(transform = "compositional") %>% tax_glom(taxrank = "Genus")
 Groups5 <- unique(as.character(meta(pseq.rel5 )$Group))

print(Groups5)
```

Write a for loop to go through each of the disease_states one by one and combine identified core taxa into a list.
```{r Imagen Core taxa}
 Groups5 <- unique(as.character(meta(pseq.rel5 )$Group))

list_core <- c() # an empty object to store information

for (n in Groups5){ # for each variable n in Group
    #print(paste0("Identifying Core Taxa for ", n))
    
    ps.sub5 <- subset_samples(pseq.rel5, Groups5 == n) 
    
    core_m <- core_members(ps.sub5, # ps.sub is phyloseq selected with only samples from g 
                           detection = 0.001, # 0.001 in at least 95% samples 
                           prevalence = 0.95)
    print(paste0("No. of core taxa in ", n, " : ", length(core_m))) 
    # print core taxa identified in each Group.
    list_core[[n]] <- core_m # add to a list core taxa for each group.
 }


# Specify colors and plot venn
# supplying colors in the order they appear in list_core
mycols <- c(ctr="#d6e2e9", `ctr+rif` ="#cbf3f0", ccl4="#fcf5c7", `ccl4+rif` = "rosybrown1") 

plot(venn(list_core),
     fills = mycols)
```

Añade: pdf("my_plot.pdf", width = 5.0, height = 6.81) 
pon el script del plot
dev.off()

# Barplots metadata 
```{r Barplots metadata }
ggplot(metadata, aes(x=Group, y=SCFA_AA)) + 
    geom_bar(stat = "identity")


#Reordenado
metadata %>%
    arrange(SCFA_AA) %>%
    mutate(Group = factor(Group, levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif"))) %>%
    ggplot( aes(x=Group, y=SCFA_AA)) + geom_bar(stat = "identity") +
    theme() +
    xlab("")


#Cambiando los tamaños
metadata %>%
    arrange(SCFA_AA) %>%
    mutate(Group = factor(Group, levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif"))) %>%
    ggplot(aes(x=Group, y=SCFA_AA)) + geom_bar(stat = "identity") +
    theme(
        axis.text.x = element_text(size = 16), axis.title.y = element_text(size = 16)) +
    xlab("")  
```

#Boxplots metadata 
```{r Boxplots metadata}
sample_data(ps_0)$Read_depth <- sample_sums(ps_0)
ps.t<- microbiome::transform(ps_0, "identity")
#Sequencing Depth
as.data.frame(meta(ps.t)) %>%
    arrange() %>%
    mutate(Group = factor(Group, levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif"))) %>%
    ggplot( aes(x=Group, y=Read_depth)) + geom_boxplot() +
    theme(axis.text.x = element_text(size = 16), axis.title.y = element_text(size = 16)) +
    xlab("")
```

#Abundance Barplots
###Phylum
```{r Barplots Phylum}
ps.rel = transform_sample_counts(ps_0, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(ps.rel, taxrank = 'Phylum', NArm = FALSE)
ps.melt <- psmelt(glom)
# change to character for easy-adjusted level
ps.melt$Phylum <- as.character(ps.melt$Phylum)
ps.melt <- ps.melt %>%
    group_by(Group, Phylum) %>%
    mutate(median=median(Abundance))

#to get the same rows together
ps.melt_sum <- ps.melt %>%
    group_by(Sample,Group,Phylum) %>%
    summarise(Abundance=sum(Abundance))%>%
    arrange(Group) %>%
    mutate(Group = factor(Group, levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif")))

ggplot(ps.melt_sum, aes(x = Sample, y = Abundance, fill = Phylum)) + 
    geom_bar(stat = "identity", aes(fill=Phylum)) + 
    labs(x="", y="%") +
    facet_wrap(~Group, scales= "free_x", nrow=1) +
    theme_classic() + 
    theme(strip.background = element_blank(), 
          axis.text.x.bottom = element_text(angle = -90))
```

###Order
```{r Barplots Order}
ps.rel = transform_sample_counts(ps_0, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(ps.rel, taxrank = 'Order', NArm = FALSE)
ps.melt <- psmelt(glom)
# change to character for easy-adjusted level
ps.melt$Order <- as.character(ps.melt$Order)
ps.melt <- ps.melt %>%
    group_by(Group, Order) %>%
    mutate(median=median(Abundance))

#to get the same rows together
ps.melt_sum <- ps.melt %>%
    group_by(Sample,Group,Order) %>%
    summarise(Abundance=sum(Abundance))%>%
    arrange(Group) %>%
    mutate(Group = factor(Group, levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif")))

ggplot(ps.melt_sum, aes(x = Sample, y = Abundance, fill = Order)) + 
    geom_bar(stat = "identity", aes(fill=Order)) + 
    labs(x="", y="%") +
    facet_wrap(~Group, scales= "free_x", nrow=1) +
    theme_classic() + 
    theme(strip.background = element_blank(), 
          axis.text.x.bottom = element_text(angle = -90))
```

###Family
```{r Barplots Family}
ps.rel = transform_sample_counts(ps_0, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(ps.rel, taxrank = 'Family', NArm = FALSE)
ps.melt <- psmelt(glom)
# change to character for easy-adjusted level
ps.melt$Family <- as.character(ps.melt$Family)
ps.melt <- ps.melt %>%
    group_by(Group, Family) %>%
    mutate(median=median(Abundance))

#to get the same rows together
ps.melt_sum <- ps.melt %>%
    group_by(Sample,Group,Family) %>%
    summarise(Abundance=sum(Abundance))%>%
    arrange(Group) %>%
    mutate(Group = factor(Group, levels=c("ctr", "ctr+rif", "ccl4", "ccl4+rif")))

ggplot(ps.melt_sum, aes(x = Sample, y = Abundance, fill = Family)) + 
    geom_bar(stat = "identity", aes(fill=Family)) + 
    labs(x="", y="%") +
    facet_wrap(~Group, scales= "free_x", nrow=1) +
    theme_classic() + 
    theme(strip.background = element_blank(), 
          axis.text.x.bottom = element_text(angle = -90))
```


